Blame and Coercion: Together Again for the Second Time

### 2019-10-15 paper.tex
missing conclusion; LCast

\newcommand{\GTLC}{\texttt{GTLC+}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\redruleS}[3]{#1 & \longrightarrow_{S(C)} & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\sOOinspect}[3]{\mathtt{Eval} \; #1 \; #2 \; #3}
\newcommand{\sOOreturn}[2]{\mathtt{Cont} \; #2 \; #1}
\newcommand{\sOOhalt}[1]{\mathtt{Halt} \; #1}
\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\mathtt{Unit}}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{unit}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{#1 \; #2}
\newcommand{\eOOcons}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\eOOcar}[1]{\mathtt{fst} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{snd} \; #1}
\newcommand{\eOOinl}[1]{\mathtt{inl} \; #1}
\newcommand{\eOOinr}[1]{\mathtt{inr} \; #1}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{unit}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \overset{#2}{\Rightarrow} #3}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{unit}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}
\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}
% Lazy D blame calculus
\newcommand{\judgeDBreduce}[2]{#1 \longrightarrow_{\mathcal{D}} #2}
\newcommand{\judgeDBreduceTrans}[2]{#1 \longrightarrow_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDBeval}[2]{#1 \Downarrow_{\mathcal{D}} #2}
\newcommand{\redrule}[3]{#1 & \longrightarrow_\mathcal{D} & #2 & #3\\}
% Normal Coercion
\newcommand{\ncProj}[2]{#1?^{#2}}
\newcommand{\ncInj}[1]{#1!}
\newcommand{\ncId}[0]{\iota}
\newcommand{\ncSeq}[2]{#1;#2}
\newcommand{\ncFail}[1]{\bot^{#1}}
\newcommand{\ncFun}[2]{\POOfun{#1}{#2}}
\newcommand{\ncProd}[2]{\POOprod{#1}{#2}}
\newcommand{\ncSum}[2]{\POOsum{#1}{#2}}
% Hyper-coercion
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}

@@

	\caption{CEK abstract machine for the \lazyD{} Blame Calculus}
	\label{machine-cekc}
\end{figure}

@@

\figref{fig:normal-coercion} defines the normal coercion. 
Let $ I $ ranges over injectable types. An injectable type is a type that can 
be cast directly to (injection) and from (projection) $ \TOOdyn $. The 
definition of injectable type depends on blame strategy. With \lazyD, all 
pre-types are injectable. With \lazyUD, only $ \POOfun{\TOOdyn}{\TOOdyn} $ 
is injectable. 
Let $ c $ ranges over coercions, a coercion is either injection, projection, 
identity, failure, sequencing, or function coercion.
Let $ \bar{c} $ ranges over wrapper coercions, which represent casts in $ 
\vOOcast{v}{c} $.
Let $ \hat{c} $ ranges over normal coercions. If we inline $ \bar{c} $ and 
re-arrange cases a little bit, the definition becomes:
\[
\begin{array}{rclr}
\stxrule{\hat{c}}{normal coercions}{
	\ncFun{\hat{c}}{\hat{c}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
	\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFail{l}}
}
\stxrulecont{
	\ncId \mid
	\ncProj{I}{l} \mid
	\ncInj{I} \mid
	\ncSeq{\ncProj{I}{l}}{\ncInj{I}} \mid
	\ncFail{l} \mid
}
\end{array}
\]

Three observations on the definition leads to hyper-coercion: 
\begin{enumerate}
	\item The length of normal coercion is at most three.
	\item Projections are always at the beginning when presented.
	\item Injections and failures are always at the end when presented.
\end{enumerate}

@@

\begin{figure}
	\[ 
	\begin{array}{lclr}
	\stxrule{c}{hyper-coercions}{
		\hyperCoercionI \mid{}
		\hyperCoercionC{h}{m}{t}
	}
	\stxrule{h}{heads}{
		\epsilon \mid{}
		?^l
	}
	\stxrule{m}{middles}{
		\POOunit \mid
		\POOfun{c_1}{c_2} \mid
		\POOprod{c_1}{c_2} \mid
		\POOsum{c_1}{c_2}
	}
	\stxrule{t}{tails}{
		\epsilon \mid{}
		! \mid{}
		\bot^l
	}
	\stxrule{\ell}{$ \mathtt{Maybe} \; l $}{
		\epsilon \mid{}
		l
	}
	\end{array}
	\]

	Shallow-consistency between middles
	\fbox{$ m \smile m $}

	%	Coercion typing \fbox{$ c : T \Longrightarrow T $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	%	\quad
	%	\inference{
	%		\typingHC{h}{T_1}{P_1} &
	%		\typingHC{m}{P_1}{P_2} &
	%		\typingHC{t}{P_2}{T_2}
	%	}{
	%		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	%	}
	%	\end{gather*}
	%	
	%	Head typing \fbox{$ \typingHC{h}{T}{P} $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\epsilon}{P}{P}}
	%	\quad
	%	\inference{}{\typingHC{?^l}{\TOOdyn}{P}}
	%	\end{gather*}
	%	
	%	Middle typing \fbox{$ \typingHC{m}{P}{P} $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\POOunit}{\POOunit}{\POOunit}}
	%	\\
	%	\inference{
	%		\typingHC{c_1}{T_3}{T_1} &
	%		\typingHC{c_2}{T_2}{T_4}
	%	}{
	%		\typingHC{\POOfun{c_1}{c_2}}{\POOfun{T_1}{T_2}}{\POOfun{T_3}{T_4}}
	%	}
	%	\\
	%	\inference{
	%		\typingHC{c_1}{T_1}{T_3} &
	%		\typingHC{c_2}{T_2}{T_4}
	%	}{
	%		
	%\typingHC{\POOprod{c_1}{c_2}}{\POOprod{T_1}{T_2}}{\POOprod{T_3}{T_4}}
	%	}
	%	\\
	%	\inference{
	%		\typingHC{c_1}{T_1}{T_3} &
	%		\typingHC{c_2}{T_2}{T_4}
	%	}{
	%		\typingHC{\POOsum{c_1}{c_2}}{\POOsum{T_1}{T_2}}{\POOsum{T_3}{T_4}}
	%	}
	%	\end{gather*}
	%	
	%	Tail typing \fbox{$ \typingHC{t}{P}{T} $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\epsilon}{P}{P}} \quad
	%	\inference{}{\typingHC{!}{P}{\TOOdyn}} \quad
	%	\inference{}{\typingHC{\bot^l}{P}{T}} \quad
	%	\end{gather*}

	\caption{Syntax of hyper-coercion}
	\label{fig:hyper-coercion}
\end{figure}

All definitions and results in this section are new.
\figref{fig:hyper-coercion} defines the syntax of hyper-coercions.
Terms, types, pre-types, blame labels, and observations are as before.

Let $ c $ ranges over hyper-coercions. We re-use the meta-variable $ c $ to 
stress that hyper-coercion is a cast representation. A hyper-coercion is either 
an identity cast between the dynamic types, or a complex including a head, a 
middle, and a tail. 
Let $ h $ ranges over heads. A head is either a no-op, or a projection 
decorated with a label, to which blame is allocated if the projection fails. 
Let $ m $ ranges over middles. There is a one-to-one 
correspondence between middles and type constructors. 
We generalize shallow-consistency to include middles in the natural way.
Let $ t $ ranges over tails. A tail is either a no-op, an injection, or a 
failure. 
Let $ \ell $ ranges over no-op and labels. It is useful when we introduce 
composition of hyper-coercions.

@@

\begin{figure}
	Composition of hyper-coercions \fbox{$ c \fatsemi^\ell c = c $}
	\[ 
	\begin{array}{rclclr}

@@

Functions that construct \lazyD{} hyper-coercions are shown in 
\figref{fig:HC-D}.

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hyper-coercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.

$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.

$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.

$ id(T) $ constructs an identity coercion of $ T $.

$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

\begin{proposition}[Hyper-coercions form a Monoid]
	For all $c : T_1 \Longrightarrow T_2$,
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c) = c$,
		\item $seq(c,id(T_2)) = c$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{proposition}

To show that hyper-coercion is a correct cast representation, we introduce an 
hyper-coercion-based machine for the \lazyD\ blame calculus, and show that it 
bi-simulates the previous machine.

The machine depends on four functions that consumes and/or construct 
hyper-coercions: $ id(T) $, $ seq(c,c) $, $ cast(T,l,T) $, and $ applyCast(T,v) 
$. The first three are described above. We will show $ applyCast $ shortly.

Some reader might have noticed that the machine is representation-independent 
w.r.t cast. That being said, if some people want to use normal coercion instead 
of hyper-coercion, they can implement the normal coercion version of the four 
functions mentioned aboved and get another abstract machine. We capture the 
idea of ``abstract cast'' by the following definition.

\subsection{Correctness of \lazyD{} Hyper-coercions}

One of our major contributions is for all cast presentation, if it is proper, 
then it is correct.

\begin{proposition}[proper cast represenations are correct]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ and $ C $ is a proper cast 
	representation
	\[
	e \Downarrow_{D,B} o \; \text{if and only if} \; 
	e \Downarrow_{D,P(C)} o
	\]
\end{proposition}

Perhaps unsurprisingly, hyper-coercion is correct.

\begin{lemma}[\lazyD{} Hyper-coercion is a proper cast representation]
	See appendix.
\end{lemma}

\begin{theorem}[\lazyD{} hyper-coercion is correct]
	If $ \judgetype{\emptyset}{e}{T} $ and $ o : T $ 
	\[
	e \Downarrow_{D,B} o \; \text{if and only if} \; 
	e \Downarrow_{D,P(H)} o
	\]
\end{theorem}

\figref{hc-applyCast} defines $ applyCast $. Applying the identity cast for the 
dynamic type succeeds immediately. When applying a compound cast, we firstly 
apply the middle, then apply the tail. We denote by $ r \; >>= \; f $ to mean 
that if $ r $ is $ \rOOsucc{v} $, the result is $ f(v) $, otherwise the result 
is the failure.
$ applyMiddle(\ell,m,v) $ and $ applyTail(t,v) $ are straightforward.

	\fbox{$ applyTail(t,v) = r $}
	\[
	\begin{array}{rclr}
	\funrule{applyTail(\bot^l,v)}{\rOOfail{l}}{}
	\funrule{applyTail(\epsilon,v)}{\rOOsucc{v}}{}
	\funrule{applyTail(!,v)}{\rOOsucc{(\hcvOOinj{P}{v})}}{v : P}
	\end{array}
	\]
	\caption{Applying hyper-coercions to values}
	\label{hc-applyCast}
\end{figure}

@@

\section{Space-efficient CEK Abstract Machine}

[move the definition of the generic machine here -Jeremy]

All definitions and results in this section are new.

We present a space-efficient CEK machine for blame calculi. This machine is
parameterized over cast representation.

@@

	Reduction \fbox{$ s \longrightarrow_{S(C)} s $}
	\[
	\begin{array}{rclr}
	& \vdots \\
	\redruleS{
		\sOOinspect{(\eOOlam{T_1}{T_2}{x}{e})}{\rho}{\kappa}
	}{
		\sOOreturn{(\hcvOOfun{id(T_1)}{\rho}{x}{e}{id(T_2)})}{\kappa}
	}{}
	\redruleS{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{ext(cast(T_1,l,T_2),\kappa)}
	}{}
	\redruleS{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{cont(\hckOOconsI{e_2}{\rho}{\kappa})}
	}{}
	\redruleS{
		\sOOreturn{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{(c,\hckOOappII{v}{\kappa})}
	}{
		\sOOinspect{e}{\rho[x:=v']}{ext(c_2,\kappa)}
	}{
		\\ & &
		\sidecond{applyCast(c_1,v) = \rOOsucc{v'}}
	}
	\redruleS{
		\sOOreturn{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{(c,\hckOOappII{v}{\kappa})}
	}{
		\sOOhalt{(\oOOblame{l})}
	}{
		\\ & &
		\sidecond{applyCast(c_1,v) = \rOOfail{l}}
	}
	\end{array}
	\]

@@

	Evaluation \fbox{$ e \Downarrow_{S(C)} o $}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{S(C)}^{*} 
		\sOOhalt{o}
	}{
		e \Downarrow_{S(C)} o
	}
	\]

	\caption{A space-efficient CKE machine for Blame Calculus.}
	\label{machine-cekcc}
\end{figure}

>>>>

\newcommand{\GTLC}{\texttt{GTLC+}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\judgeType}[2]{#1 : #2}
\newcommand{\judgeTypeFT}[3]{#1 : #2 \Longrightarrow #3} % FT = From To
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\sOOinspect}[3]{\mathtt{Eval} \; #1 \; #2 \; #3}
\newcommand{\sOOreturn}[2]{\mathtt{Cont} \; #2 \; #1}
\newcommand{\sOOhalt}[1]{\mathtt{Halt} \; #1}
\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\mathtt{Unit}}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{unit}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{#1 \; #2}
\newcommand{\eOOcons}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\eOOcar}[1]{\mathtt{fst} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{snd} \; #1}
\newcommand{\eOOinl}[1]{\mathtt{inl} \; #1}
\newcommand{\eOOinr}[1]{\mathtt{inr} \; #1}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{unit}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \overset{#2}{\Rightarrow} #3}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{unit}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}
\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}
% Lazy D blame calculus on space-inefficient CEK
\newcommand{\judgeDBreduce}[2]{#1 \longrightarrow_{\mathcal{D}} #2}
\newcommand{\judgeDBreduceTrans}[2]{#1 \longrightarrow_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDBeval}[2]{#1 \Downarrow_{\mathcal{D}} #2}
\newcommand{\redrule}[3]{#1 & \longrightarrow_\mathcal{D} & #2 & #3\\}
% blame calculus on space-efficient CEK
\newcommand{\judgeSreduce}[3]{#2 \longrightarrow_{\mathcal{S}(#1)} #3}
\newcommand{\judgeSreduceTrans}[3]{#2 \longrightarrow_{\mathcal{S}(#1)}^{*} #3}
\newcommand{\judgeSeval}[3]{#2 \Downarrow_{\mathcal{S}(#1)} #3}
\newcommand{\redruleS}[3]{#1 & \longrightarrow_{\mathcal{S}(C)} & #2 & #3\\}
% Normal Coercion
\newcommand{\ncProj}[2]{#1?^{#2}}
\newcommand{\ncInj}[1]{#1!}
\newcommand{\ncId}[0]{\iota}
\newcommand{\ncSeq}[2]{#1;#2}
\newcommand{\ncFail}[1]{\bot^{#1}}
\newcommand{\ncFun}[2]{\POOfun{#1}{#2}}
\newcommand{\ncProd}[2]{\POOprod{#1}{#2}}
\newcommand{\ncSum}[2]{\POOsum{#1}{#2}}
% Hyper-coercion
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}

@@

	\caption{CEK abstract machine ($ \mathcal{D} $) for the \lazyD{} Blame 
	Calculus}
	\label{machine-cekc}
\end{figure}

@@

\figref{fig:normal-coercion} defines the normal coercion. 
Types, pre-types, and blame labels are as before.
Let $ I $ ranges over injectable types. An injectable type is a type that can 
be cast directly to (injection) and from (projection) $ \TOOdyn $. The 
definition of injectable type depends on blame strategy. With \lazyD, all 
pre-types are injectable. With \lazyUD, only $ \POOfun{\TOOdyn}{\TOOdyn} $ 
is injectable. 
Let $ c $ ranges over coercions, a coercion is either injection, projection, 
identity, failure, sequencing, or function coercion.
Let $ \bar{c} $ ranges over wrapper coercions, which represent casts in $ 
\vOOcast{v}{c} $.
Let $ \hat{c} $ ranges over normal coercions. If we inline $ \bar{c} $ and 
re-arrange cases a little bit, the definition becomes:
\[
\begin{array}{rclr}
\stxrule{\hat{c}}{normal coercions}{
	\ncFun{\hat{c}}{\hat{c}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
	\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFail{l}}
}
\stxrulecont{
	\ncId \mid
	\ncProj{I}{l} \mid
	\ncInj{I} \mid
	\ncSeq{\ncProj{I}{l}}{\ncInj{I}} \mid
	\ncFail{l} \mid
}
\end{array}
\]

Three observations on this definition leads to hyper-coercion: 
\begin{enumerate}
	\item The length of normal coercion is at most three.
	\item Projections are always at the beginning when presented.
	\item Injections and failures are always at the end when presented.
\end{enumerate}

@@

\begin{figure}
	Syntax
	\[
	\begin{array}{lclr}
	\stxrule{I}{injectable types (\lazyD)}{P}
	\stxrule{I}{injectable types (\lazyUD)}{
		\POOunit \mid
		\POOfun{\TOOdyn}{\TOOdyn} \mid
		\POOprod{\TOOdyn}{\TOOdyn} \mid
		\POOsum{\TOOdyn}{\TOOdyn}
	}
	\stxrule{c}{hyper-coercions}{
		\hyperCoercionI \mid{}
		\hyperCoercionC{h}{m}{t}
	}
	\stxrule{h}{heads}{
		\epsilon \mid{}
		?^l
	}
	\stxrule{m}{middles}{
		\POOunit \mid
		\POOfun{c_1}{c_2} \mid
		\POOprod{c_1}{c_2} \mid
		\POOsum{c_1}{c_2}
	}
	\stxrule{t}{tails}{
		\epsilon \mid{}
		! \mid{}
		\bot^l
	}
	\end{array}
	\]

	Hyper-coercion typing \fbox{$ c : T \Longrightarrow T $}
	\begin{gather*}
	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	\quad
	\inference{
		\typingHC{h}{T_1}{P_1} &
		\typingHC{m}{P_1}{P_2} &
		\typingHC{t}{P_2}{T_2}
	}{
		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	}
	\end{gather*}

	Head typing \fbox{$ \typingHC{h}{T}{P} $}
	\begin{gather*}
	\inference{}{\typingHC{\epsilon}{P}{P}}
	\quad
	\inference{}{\typingHC{?^l}{\TOOdyn}{I}}
	\end{gather*}

	Middle typing \fbox{$ \typingHC{m}{T}{T} $}
	\begin{gather*}
	\inference{}{\typingHC{\POOunit}{\POOunit}{\POOunit}}
	\quad
	\inference{
		\typingHC{c_1}{T_3}{T_1} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOfun{c_1}{c_2}}{\POOfun{T_1}{T_2}}{\POOfun{T_3}{T_4}}
	}
	\\
	\inference{
		\typingHC{c_1}{T_1}{T_3} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOprod{c_1}{c_2}}{\POOprod{T_1}{T_2}}{\POOprod{T_3}{T_4}}
	}
	\quad
	\inference{
		\typingHC{c_1}{T_1}{T_3} &
		\typingHC{c_2}{T_2}{T_4}
	}{
		\typingHC{\POOsum{c_1}{c_2}}{\POOsum{T_1}{T_2}}{\POOsum{T_3}{T_4}}
	}
		\end{gather*}

		Tail typing \fbox{$ \typingHC{t}{P}{T} $}
		\begin{gather*}
		\inference{}{\typingHC{\epsilon}{P}{P}} \quad
		\inference{}{\typingHC{!}{I}{\TOOdyn}} \quad
		\inference{}{\typingHC{\bot^l}{P}{T}} \quad
		\end{gather*}

	\caption{Definition of Hyper-coercion (HC)}
	\label{fig:hyper-coercion}
\end{figure}

This section presents our first contribution, hyper-coercion (HC). 
\figref{fig:hyper-coercion} defines the syntax of HC.
Types, pre-types, and blame labels are as before.

Let $ c $ ranges over hyper-coercions. A hyper-coercion is either 
an identity cast between the dynamic types, or a complex including a head, a 
middle, and a tail. 
Let $ h $ ranges over heads. A head is either a no-op, or a projection.
Let $ m $ ranges over middles. There is a one-to-one 
correspondence between middles and type constructors. 
We generalize shallow-consistency to middles $ m \smile m $ in the natural way.
Let $ t $ ranges over tails. A tail is either a no-op, an injection, or a 
failure. 
Let $ \ell $ ranges over no-op and labels. It is useful when we introduce 
composition of hyper-coercions.

@@

\begin{figure}
	\[
	\begin{array}{rclr}
	\stxrule{\ell}{Maybe $ l $}{\epsilon \mid l}
	\end{array}
	\]

	Composition of hyper-coercions \fbox{$ c \fatsemi^\ell c = c $}
	\[ 
	\begin{array}{rclclr}

@@

\lazyD{} HC constructors are in \figref{fig:HC-D}.

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hyper-coercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

\begin{proposition}[\lazyD\ Hyper-coercions is a Monoid]
	For all $c : T_1 \Longrightarrow T_2$,
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c) = c$,
		\item $seq(c,id(T_2)) = c$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{proposition}

The correctness proof of \lazyD\ HC is deferred to \secref{sec:LDHC-correct} 
because it relies on the space-efficient CEK abstract machine in 
\secref{sec:CEKS}.

@@

\section{Space-efficient CEK Abstract Machine}
\label{sec:CEKS}

This section presents our second contribution, a space-efficient CEK abstract 
machine. This machine is parameterized over \emph{cast representation} (CR).

@@

	Reduction \fbox{$ \judgeSreduce{C}{s}{s} $}
	\[
	\begin{array}{rclr}
	& \vdots \\
	\redruleS{
		\sOOinspect{(\eOOlam{T_1}{T_2}{x}{e})}{\rho}{\kappa}
	}{
		\sOOreturn{(\hcvOOfun{id(T_1)}{\rho}{x}{e}{id(T_2)})}{\kappa}
	}{}
	\redruleS{
		\sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\rho}{\kappa}
	}{
		\sOOinspect{e}{\rho}{ext(cast(T_1,l,T_2),\kappa)}
	}{}
	\redruleS{
		\sOOinspect{(\eOOcons{e_1}{e_2})}{\rho}{\kappa}
	}{
		\sOOinspect{e_1}{\rho}{cont(\hckOOconsI{e_2}{\rho}{\kappa})}
	}{}
	\redruleS{
		\sOOreturn{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{(c,\hckOOappII{v}{\kappa})}
	}{
		\sOOinspect{e}{\rho[x:=v']}{ext(c_2,\kappa)}
	}{
		\\ & &
		\sidecond{applyCast(c_1,v) = \rOOsucc{v'}}
	}
	\redruleS{
		\sOOreturn{(\hcvOOfun{c_1}{\rho}{x}{e}{c_2})}{(c,\hckOOappII{v}{\kappa})}
	}{
		\sOOhalt{(\oOOblame{l})}
	}{
		\\ & &
		\sidecond{applyCast(c_1,v) = \rOOfail{l}}
	}
	\end{array}
	\]

@@

	Evaluation \fbox{$ \judgeSeval{C}{e}{o} $}
	\[
	\inference{
		\judgeSreduceTrans{C}{
			\sOOinspect{e}{\emptyset}{cont(\hckOOmt)}
		}{
			\sOOhalt{o}
		}		
	}{
		\judgeSeval{C}{e}{o}
	}
	\]

	\caption{Space-efficient CEK machine $ \mathcal{S}(C) $}
	\label{machine-cekcc}
\end{figure}

\figref{machine-cekcc} defines the machine. 
Let $ v $ ranges over value. A value is either the unit value, function, pair, 
left injection of sum, right injection of sum, and injection to $ \TOOdyn $.
Cast results ($ r $) and machine states ($ s $) are as before. 
Let $ \kappa $ ranges over continuations and let $ k $ ranges over 
pre-continuations. 
A continuation is now a pair where the first part is a cast and the second 
part is a pre-continuation. 
Pre-continuation are like the continuations before.
